---
title: "Reproducible Research Peer Assignment 1"
author: "Ray Hookway"
date: "October 17, 2015"
output: html_document
---

##Loading and preprocessing the data

```{r}
library(dplyr)
rawdata <- read.csv('activity.csv')
data <- filter(rawdata, !is.na(steps))
```

##What is mean total number of steps taken per day?
###Total steps taken per day

```{r}
steps <- group_by(data, date) %>%
    summarise(
        totalSteps = sum(steps),
        meanSteps = mean(steps),
        medianSteps = median(steps)
    )
print.data.frame(select(steps, date, totalSteps))
print(mean(steps$totalSteps))
print(median(steps$totalSteps))
```

###Histogram of total steps taken each day.

```{r}
hist(steps$totalSteps, main = 'Histogram of steps', xlab = 'steps per day', breaks = 30)
```

###Mean and median of total steps taken per day.

```{r}
print(as.data.frame(steps))
```

##What is the average daily activity pattern?
###Time series plot

```{r}
timeseries <- group_by(data, interval) %>% 
  summarise(totalSteps = sum(steps), meanSteps = mean(steps), medianSteps = median(steps))
naSteps <- sum(timeseries$meanSteps)
plot(timeseries$interval, timeseries$totalSteps, type='l', main = 'Time series', xlab = 'interval', ylab = 'steps')
```

###Interval containing the maximum number of steps

```{r}
print(timeseries[which.max(timeseries$totalSteps),])
```

##Impute missing values

###Number of missing values

```{r}
sum(is.na(rawdata$steps))
```

###Strategy for imputing missing values of steps

Imputed missing value is the average value for the interval across all days.

###Dataset with imputed values

```{r}
mutated <- left_join(rawdata, timeseries, by = 'interval') %>%
  mutate(steps = ifelse(!is.na(steps), steps, meanSteps)) %>%
  select(steps, date, interval)
head(mutated)
```

###Histogram of total number of steps taken

```{r}
mutatedSteps <- group_by(mutated, date) %>%
    summarise(
        totalSteps = sum(steps),
        meanSteps = mean(steps),
        medianSteps = median(steps)
    )
hist(mutatedSteps$totalSteps, main = 'Histogram of steps per day (mutated data)', xlab = 'steps per day', breaks = 30)
```

Values differ from earlier, non-mutated, data. The primary difference is to add value to the bucket between 10000 and
15000. This makes sense since the the sum of the mean values of the number of steps per day is `r naSteps`.

###Mean and median number of steps taken per day (mutated data)

```{r}
print.data.frame(mutatedSteps)
```

##Differences in activity between weekends and weekdays

```{r}
mutated$dow <- as.factor(ifelse(grepl('Sat|Sun', weekdays(as.Date(mutated$date))), 'weekend', 'weekday'))
head(mutated)
x <- group_by(mutated, dow, interval) %>% summarise(meanSteps = mean(steps))
library(lattice)
xyplot(meanSteps ~ interval | dow, x, type = 'l', layout = c(1,2), main = 'Steps per interval')
```
