---
title: "Reproducible Research Peer Assignment 1"
author: "Ray Hookway"
date: "October 17, 2015"
output: html_document
---

##Loading and preprocessing the data

```{r}
library(dplyr)
rawdata <- read.csv('activity.csv')
data <- filter(rawdata, !is.na(steps))
```

##What is mean total number of steps taken per day?
###Total steps taken per day

```{r}
steps <- group_by(data, date) %>%
    summarise(
        totalSteps = sum(steps),
        meanSteps = mean(steps),
        medianSteps = median(steps)
    )
print.data.frame(select(steps, date, totalSteps))
```

###Histogram of total steps taken each day.

```{r}
hist(steps$totalSteps)
```

###Mean and median of total steps taken per day.

```{r}
print(as.data.frame(steps))
```

##What is the average daily activity pattern?
###Time series plot

```{r}
timeseries <- group_by(data, interval) %>% 
  summarise(totalSteps = sum(steps), meanSteps = mean(steps), medianSteps = median(steps))
plot(timeseries$interval, timeseries$totalSteps, type='l')
```

###Interval containing the maximum number of steps

```{r}
print(timeseries[which.max(timeseries$totalSteps),])
```

##Impute missing values

###Number of missing values

```{r}
sum(is.na(rawdata$steps))
```

###Strategy for imputing missing values of steps

Imputed missing value is the average value for the interval across all days.

###Dataset with imputed values

```{r}
mutated <- left_join(rawdata, timeseries, by = 'interval') %>%
  mutate(steps = ifelse(!is.na(steps), steps, meanSteps)) %>%
  select(steps, date, interval)
head(mutated)
```

###Histogram of total number of steps taken

```{r}
mutatedSteps <- group_by(mutated, date) %>%
    summarise(
        totalSteps = sum(steps),
        meanSteps = mean(steps),
        medianSteps = median(steps)
    )
hist(mutatedSteps$totalSteps)
```

###Mean and median number of steps taken per day (mutated data)

```{r}
print.data.frame(mutatedSteps)
```


